A Heuristic Approach towards Drawings of Graphs with High Crossing Resolution

### 2018-05-08 paper/paper.tex
Working a bit more one the algorithm; finished the description of the main part.

\documentclass{llncs}
\usepackage{amsmath,amssymb}
\usepackage{graphicx,rotating}
\usepackage[font=small]{subfig,caption}
\usepackage{algorithm2e}
\usepackage{todonotes,lineno,paralist}
\usepackage{placeins,rotating,hyperref}
\graphicspath{{figures/}}

\author{Michael~A.~Bekos, Henry~F\"orster, Christian-Marius~Geckeler, Michael~Kaufmann, Amad\"aus~Spallek, Jan~Splett}
\title{A Heuristic Approach towards Drawings of Graphs with High Crossing Resolution}

\institute{
Wilhelm-Schickhard-Institut f\"ur Informatik, Universit\"at T\"ubingen, Germany\\
\texttt{\{bekos,foersth,mk\}@informatik.uni-tuebingen.de}\\
\texttt{\{christian-marius.geckeler,amadaeus.spallek,jan.splett\}@student.uni-tuebingen.de}
}

@@

Given a drawing $\Gamma(G)$ of $G$ on the plane, we denote by $p_u=(x_u,y_u)$ the position of vertex $u \in V$ of $G$ in $\Gamma(G)$. For a pair of vertices $u,v\in V$ of $G$, the unit-length vector from $p_u$ to $p_v$ is denoted as $\overrightarrow{p_up_v}$, while the line segment delimited by $p_u$ and $p_v$ as $\overline{p_up_v}$.

@@

So far, we have described the main idea of our iterative algorithm, which at each iteration step chooses uniformly at random a vertex from the vertex-pool of the current drawing to move, so to improve the crossing resolution. We have not described, however, how the chosen vertex is moved, i.e., how to compute its new position. To this end, let $v_i$ be the vertex that has been chosen from the vertex-pool of the drawing $\Gamma_{i-1}$ at the $i$-th iteration of our algorithm.

\begin{itemize}
\item Description of the randomized algorithm.
\item Big graphs $\rightarrow$ focus on critical edges.
\item Mimics the human. 
\item Avoiding local minima.
\item Crossing calculation. 
\end{itemize}

@@

>>>>

\documentclass{llncs}
\usepackage{amsmath,amssymb}
\usepackage{graphicx,rotating}
\usepackage[font=small]{subfig,caption}
\usepackage{algorithm2e}
\usepackage{todonotes,lineno,paralist,soul}
\usepackage{placeins,rotating,hyperref}
\graphicspath{{figures/}}

\author{Michael~A.~Bekos, Henry~F\"orster, Christian-Marius~Geckeler, Lukas Holl\"ander, Michael~Kaufmann, Amad\"aus~Spallek, Jan~Splett}
\title{A Heuristic Approach towards Drawings of Graphs with High Crossing Resolution}

\institute{
Wilhelm-Schickhard-Institut f\"ur Informatik, Universit\"at T\"ubingen, Germany\\
\texttt{\{bekos,foersth,mk\}@informatik.uni-tuebingen.de}\\
\texttt{\{christian-marius.geckeler,jan-lukas.hollaender,amadaeus.spallek,jan.splett\} @student.uni-tuebingen.de}
}

@@

Given a drawing $\Gamma(G)$ of $G$ on the plane, we denote by $p(u)=(x_u,y_u)$ the position of vertex $u \in V$ of $G$ in $\Gamma(G)$. For a pair of vertices $u,v\in V$ of $G$, the unit-length vector from $p(u)$ to $p(v)$ is denoted as $\overrightarrow{p(u)p(v)}$, while the line segment delimited by $p(u)$ and $p(v)$ as $\overline{p(u)p(v)}$.

@@

So far, we have described the main idea of our iterative algorithm, which at each iteration step chooses uniformly at random a vertex from the vertex-pool of the current drawing to move, so to improve the crossing resolution. We have not described, however, how the chosen vertex is moved, i.e., how we compute its new position in the next drawing.

Let $v_i$ be the vertex that has been chosen from the vertex-pool of the drawing $\Gamma_{i-1}$ at the $i$-th iteration of our algorithm. Recall that the crossing resolution $c(\Gamma_{i})$ of drawing $\Gamma_{i}$ that is obtained after the $i$-th iteration of the algorithm must be at least as large as the crossing resolution $c(\Gamma_{i-1})$ of drawing $\Gamma_{i-1}$ (by the invariant of our algorithm), that is, $c(\Gamma_i) \ge c(\Gamma_{i-1})$. To compute the position of vertex $v_i$ in drawing $\Gamma_i$, we consider a set of $\rho$ rays $r_0,r_1,\ldots,r_{\rho-1}$ that all emanate from the point $p(v_i)$ of vertex $v_i$ in drawing $\Gamma_{i-1}$, such that the angle formed by ray $r_j$, with $j=0,1,\ldots,\rho-1$, and the horizontal axis equals to $2j\pi/\rho$, where $\rho>0$ is an integer parameter of the algorithm. These rays are then rotated by an angle that is chosen uniformly at random in the interval $[0,2\pi]$. The position of vertex $v_i$ in drawing $\Gamma_i$ will eventually be along one of the rays $r_0,r_1,\ldots,r_{\rho-1}$. More precisely, for each ray $r_i$ we chose a distance value $\delta_i$ uniformly at random from the interval $[0,\delta]$, where $\delta$ is also a parameter of the algorithm. For each $j=0,1,\ldots,\rho-1$, a new point $\pi_j$ on the plane is obtained translating $p(u)$ along ray $r_j$ at a distance $\delta_j$; we say that point $\pi_j$ is \emph{feasible}, if the crossing resolution of the drawing obtained by placing vertex $v_i$ at point $\pi_j$ and by keeping all other vertices of $G$ in the same positions as in drawing $\Gamma_{i-1}$ is at least as large as the crossing resolution of drawing $\Gamma_{i-1}$\todo{Add a figure illustrating the approach.}.

If none of the points $\pi_j$, with $j=0,1,\ldots,\rho-1$ is feasible, then the new position of vertex $v_i$ in drawing $\Gamma_i$ is $p(v_i)$, that is, the same as in drawing $\Gamma_{i-1}$, since the crossing resolution of drawing $\Gamma_i$ must be at least as large as the one of $\Gamma_{i-1}$. On the other hand, if there is one or more feasible points (among $\pi_0,\pi_1,\ldots,\pi_{\rho-1}$), then one may consider two different approaches to determine the new position of vertex $v_i$ in drawing $\Gamma_i$. The most natural is to chose this feasible point that maximizes the crossing resolution of the obtained drawing. As an alternative, one may rely again on randomization and chose uniformly at random one the feasible points as the new position of vertex $v_i$ in drawing $\Gamma_i$. In our implementation, we did not observe any significant difference in these two approaches (in terms of the crossing resolution of the obtained drawings), so we simply adopted the first one. In Section~\ref{ssec:termination}, we describe conditions  The iteration steps

\begin{itemize}
\item \st{Description of the randomized algorithm}.
\item \st{Big graphs $\rightarrow$ focus on critical edges}.
\item Avoiding local minima.
\item Crossing calculation. 
\end{itemize}

@@

From the above description, it is not difficult to see that our algorithm in a sense mimics the way a human would try to increase the crossing resolution of a drawing; first, one would try to identify...

\subsection{Conditions for termination.}
\label{ssec:termination}
